Proportions of extreme values the usual direction

### 2021-04-14 analysis/reports/submission2/manuscript_main_rev.Rmd
maybe?

<!-- ```{r, fig.dim = c(4, 4)} -->
<!-- ggplot(all_di, aes(sim_pos_from_best, real_po, color = `Observed \npercentile score`)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_point(data = filter(all_di, s0 < 0)) + -->
<!--   geom_abline(slope = 1, intercept = 0) + -->
<!--   scale_color_viridis_d(end = .8, direction = -1) + -->
<!--   xlab("Mean of elements from feasible set") + -->
<!--   ylab("Observed value") + -->
<!--   ggtitle("Dissimilarity to the central tendency") +  theme(legend.position = "bottom")#, -->
<!-- # -->
<!-- # -->
<!-- # ggplot(filter(all_di, nparts < 10 ^ 50), aes(nparts, sim_pos_from_best, color = Dataset)) + -->
<!-- #   geom_point(alpha = .3) + -->
<!-- #   geom_point(data = filter(all_di, s0 < 0)) + -->
<!-- #   scale_color_viridis_d(end = .9) + -->
<!-- #   xlab("Number of elements in the feasible set") + -->
<!-- #   ylab("Mean dissimilarity of feasible set to central tendency") + -->
<!-- #   scale_x_log10() + -->
<!-- #   ggtitle("Over the size of the feasible set") -->
<!-- # ), ncol = 1, -->
<!-- #top = textGrob("Dissimilarity to central tendency")) -->

<!-- ``` -->

@@

}
#
# gridExtra::grid.arrange(grobs = list(
#   plot_narrowness(all_di, "sim_pos_from_best", "Dissimilarity to central tendency"),
#   plot_narrowness_legend(all_di)),
#   ncol = 1)
```

@@

<!--   plot_narrowness_hist(all_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_narrowness_hist(all_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(all_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white")) -->

<!-- ``` -->

<!-- # Gentry -->

<!-- ```{r} -->

<!-- gentry_di <- filter(all_di, Dataset == "Gentry") -->

<!-- gentry_di <- gentry_di %>% -->
<!--   mutate(low_avg_abund = ifelse(n0/s0 < 3, "N:S < 3", "N:S > 3")) -->

<!--   plot_percentile_hist(gentry_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "low_avg_abund"), -->
<!--   plot_percentile_hist(gentry_di, "skew_percentile_mean", "\nSkewness", facetvar = "low_avg_abund", min_s0 = 3), -->
<!--   plot_percentile_hist(gentry_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "low_avg_abund"), -->
<!--   plot_percentile_hist(gentry_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "low_avg_abund") -->
<!-- ), ncol = 2, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->

<!-- ``` -->

<!-- # Comparison of FIA and comparably sized communities (Figure 4) -->

<!-- ```{r, fig.dim = c(10, 5)} -->

<!-- fia_max_s = max(filter(all_di, dat == "fia")$s0) -->
<!-- fia_max_n = max(filter(all_di, dat == "fia")$n0) -->

<!-- fia = filter(all_di, dat == "fia") -->

<!-- small_di <- all_di %>% -->
<!--   filter(s0 <= fia_max_s, -->
<!--          n0 <= fia_max_n) %>% -->
<!--   mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"), -->
<!--          right_corner = FALSE) -->

<!-- for(i in 1:nrow(small_di)) { -->
<!--   if(small_di$dat[i] != "fia") { -->
<!--     fia_match_s0 = filter(fia, s0 >= small_di$s0[i]) -->

<!--     max_n0_for_this_s0 = max(fia_match_s0$n0) -->

<!--     small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0 -->

<!--   } -->
<!-- } -->

<!-- small_di <- small_di %>% -->
<!--   filter(!right_corner) -->
<!-- not_fia <- filter(small_di, fia_yn != "fia") -->
<!-- fia = filter(small_di, dat == "fia") -->

<!-- small_di_s_n <- small_di %>% -->
<!--   select(s0, n0) %>% -->
<!--   distinct() %>% -->
<!--   mutate(in_fia = FALSE, -->
<!--          in_not_fia = FALSE) -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->
<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 -->
<!--   small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0 -->

<!-- } -->

<!-- small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia) -->

<!-- set.seed(1977) -->

<!-- subsamples <- list() -->

<!-- for(i in 1:nrow(small_di_s_n)) { -->

<!--   this_s0 = small_di_s_n$s0[i] -->
<!--   this_n0 = small_di_s_n$n0[i] -->

<!--   fia_matches = filter(fia, s0 == this_s0, n0 == this_n0) -->

<!--   not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0) -->

<!--   n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches)) -->

<!--   fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F) -->
<!--   not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F) -->

<!--   subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ])) -->

<!-- } -->

<!-- sub_di <- bind_rows(subsamples) %>% -->
<!--   mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets")) -->

<!-- fia_di <- filter(sub_di, fia_yn == "fia") -->
<!-- other_di <- filter(sub_di, fia_yn != "fia") -->

<!-- ks_compare <- function(fia_df, other_df, compare_var) { -->

<!--   if(grepl("skew", compare_var)) { -->
<!--     fia_df <- filter(fia_df, s0 > 2) -->
<!--     other_df <- filter(fia_df, s0 >2) -->
<!--   } -->

<!--   if(grepl("sim_pos", compare_var)) { -->
<!--     fia_df <- filter(fia_df, nparts > 20) -->
<!--     other_df <- filter(other_df, nparts > 20) -->
<!--   } else { -->
<!--       fia_df <- filter(fia_df, nparts > 40) -->
<!--     other_df <- filter(other_df, nparts > 40) -->
<!--   } -->

<!--   kstest <- ks.test(fia_df[[compare_var]], other_df[[compare_var]]) -->

<!--   return(data.frame( -->
<!--     var = compare_var, -->
<!--     d = kstest$statistic, -->
<!--     p = kstest$p.value -->
<!--   )) -->
<!-- } -->

<!-- simpson_rat_ks <- ks_compare(fia_di, other_di, "simpson_95_ratio_2t") -->
<!-- skew_rat_ks <- ks_compare(fia_di, other_di, "skew_95_ratio_2t") -->
<!-- shannon_rat_ks <- ks_compare(fia_di, other_di, "shannon_95_ratio_2t") -->
<!-- nsingletons_rat_ks <- ks_compare(fia_di, other_di, "nsingletons_95_ratio_2t") -->
<!-- diss_rat_ks <- ks_compare(fia_di, other_di, "sim_pos_from_best") -->

<!-- bind_rows(simpson_rat_ks, skew_rat_ks, shannon_rat_ks, nsingletons_rat_ks, diss_rat_ks) -->

<!-- simpson_perc_ks <- ks_compare(fia_di, other_di, "simpson_percentile") -->
<!-- skew_perc_ks <- ks_compare(fia_di, other_di, "skew_percentile_excl") -->
<!-- shannon_perc_ks <- ks_compare(fia_di, other_di, "shannon_percentile") -->
<!-- nsingletons_perc_ks <- ks_compare(fia_di, other_di, "nsingletons_percentile_excl") -->
<!-- diss_perc_ks <- ks_compare(fia_di, other_di, "real_po_percentile_excl") -->

<!-- bind_rows(simpson_perc_ks, skew_perc_ks, shannon_perc_ks, nsingletons_perc_ks, diss_perc_ks) -->

<!--   plot_percentile_hist(sub_di, "real_po_percentile_mean", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset", tails = 1), -->
<!--   plot_percentile_hist(sub_di, "nsingletons_percentile_mean", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(sub_di, "skew_percentile_mean", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_percentile_hist(sub_di, "simpson_percentile_mean", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_percentile_hist(sub_di, "shannon_percentile_mean", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!-- top = textGrob("Figure 2", gp = gpar(fill = "white")), left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), -->
<!-- bottom = textGrob("Percentile rank of observed value relative to feasible set"), gp = gpar(fill = "white")) -->

<!--   plot_narrowness_hist(sub_di, "sim_pos_from_best", "Dissimilarity to the \ncentral tendency", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "nsingletons_95_ratio_2t", "\nNumber of rare species", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "skew_95_ratio_2t", "\nSkewness", facetvar = "Dataset", min_s0 = 3), -->
<!--   plot_narrowness_hist(sub_di, "simpson_95_ratio_2t", "\nSimpson evenness", facetvar = "Dataset"), -->
<!--   plot_narrowness_hist(sub_di, "shannon_95_ratio_2t", "\nShannon diversity", facetvar = "Dataset") -->
<!-- ), ncol = 5, -->
<!--  left = textGrob("Number of communities", rot = 90, gp = gpar(fill = "black")), gp = gpar(fill = "white")) -->

<!-- diss_results <- sub_di %>% -->
<!--   filter(nparts > 20) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_diss = mean(real_po_percentile_excl > 95), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- skew_results <- sub_di %>% -->
<!--   filter(nparts > 40, s0 > 2) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_skew = mean(skew_percentile_excl > 97.5), -->
<!--             low_skew = mean(skew_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites), -->
<!--          `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- metric_results <- sub_di %>% -->
<!--   filter(nparts > 40) %>% -->
<!--   group_by(Dataset) %>% -->
<!--   summarize(high_simpson = mean(simpson_percentile_excl > 97.5), -->
<!--             low_simpson = mean(simpson_percentile < 2.5), -->
<!--             high_shannon = mean(shannon_percentile_excl > 97.5), -->
<!--             low_shannon = mean(shannon_percentile < 2.5), -->
<!--             high_nsingletons = mean(nsingletons_percentile_excl > 97.5), -->
<!--             low_nsingletons = mean(nsingletons_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites), -->
<!--          `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites), -->
<!--          `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites), -->
<!--          `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites), -->
<!--          `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites), -->
<!--          `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- all_results <- left_join(diss_results, skew_results) %>% -->
<!--   left_join(metric_results) -->

<!-- cols1 <- c("Dataset", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon") -->

<!-- all_results %>% -->
<!--   select(cols1) -->

<!-- ``` -->

@@

gridExtra::grid.arrange(grobs = list(nhist, phist), nrow = 2, top = textGrob("Dissimilarity to the central tendency"))

# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "skew_95_ratio_2t", "Skewness") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "skew_percentile_mean", "Skewness") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)

# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_95_ratio_2t", "Number of rare species") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "nsingletons_percentile_mean", "Number of rare species") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_95_ratio_2t", "Simpson") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "simpson_percentile_mean", "Simpson") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# 
# plot_narrowness_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_95_ratio_2t", "Shannon") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
# 
# plot_percentile_hist(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), "shannon_percentile_mean", "Shannon") +
#   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y", nrow = 2, ncol = 3)
```

<!-- ```{r, fig.dim = c(8,3)} -->
<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund")), aes(s0, n0, color = `Number of elements`)) + -->
<!--   geom_point(alpha = .1) + -->
<!--   xlim(0, 30) + -->
<!--   ylim(0,150) + -->
<!--     facet_wrap(vars(Dataset), nrow = 1) -->
<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts > 2000, nparts < 1e+20), aes(nparts)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y") + -->
<!--   scale_x_log10() -->

<!-- ggplot(filter(all_di, dat %in% c("fia", "mcdb", "misc_abund"), nparts < 2000), aes(nparts)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`, Dataset), scales = "free_y") + -->
<!--   scale_x_log10() -->

<!-- ``` -->

<!-- ```{r} -->

<!-- all_di <- all_di %>% -->
<!--   mutate( `Number of elements` = ifelse(nparts < 1000, "Less than 1000", "More than 1000")) -->
<!-- diss_results <- all_di %>% -->
<!--   filter(nparts > 20) %>% -->
<!--   group_by(Dataset,  `Number of elements`) %>% -->
<!--   summarize(high_diss = mean(real_po_percentile_excl > 95), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High dissimilarity` = paste0(signif(100 * high_diss, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->
<!-- ggplot(filter(all_di, dat == "mcdb"), aes(skew_95_ratio_2t)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(vars(`Number of elements`), scales = "free_y") -->

<!-- skew_results <- all_di %>% -->
<!--   filter(nparts > 40, s0 > 2) %>% -->
<!--   group_by(Dataset,  `Number of elements`) %>% -->
<!--   summarize(high_skew = mean(skew_percentile_excl > 97.5), -->
<!--             low_skew = mean(skew_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High skew` = paste0(signif(100 * high_skew, 2), "%; n = ", nsites), -->
<!--          `Low skew` = paste0(signif(100 * low_skew, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- metric_results <- all_di %>% -->
<!--   filter(nparts > 40) %>% -->
<!--   group_by(Dataset, `Number of elements`) %>% -->
<!--   summarize(high_simpson = mean(simpson_percentile_excl > 97.5), -->
<!--             low_simpson = mean(simpson_percentile < 2.5), -->
<!--             high_shannon = mean(shannon_percentile_excl > 97.5), -->
<!--             low_shannon = mean(shannon_percentile < 2.5), -->
<!--             high_nsingletons = mean(nsingletons_percentile_excl > 97.5), -->
<!--             low_nsingletons = mean(nsingletons_percentile < 2.5), -->
<!--             nsites = dplyr::n()) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(`High Simpson` = paste0(signif(100 * high_simpson, 2), "%; n = ", nsites), -->
<!--          `Low Simpson` = paste0(signif(100 * low_simpson, 2), "%; n = ", nsites), -->
<!--          `High Shannon` = paste0(signif(100 * high_shannon, 2), "%; n = ", nsites), -->
<!--          `Low Shannon` = paste0(signif(100 * low_shannon, 2), "%; n = ", nsites), -->
<!--          `High proportion of rare species` = paste0(signif(100 * high_nsingletons, 2), "%; n = ", nsites), -->
<!--          `Low proportion of rare species` = paste0(signif(100 * low_nsingletons, 2), "%; n = ", nsites)) %>% -->
<!--   select(-nsites) -->

<!-- all_results <- left_join(diss_results, skew_results) %>% -->
<!--   left_join(metric_results) -->

<!-- cols1 <- c("Dataset", "Number of elements", "High dissimilarity", "High proportion of rare species", "High skew", "Low Simpson", "Low Shannon") -->

<!-- all_results %>% -->
<!--   select(cols1) -->

<!-- ``` -->

>>>>

@@

@@

@@

gridExtra::grid.arrange(grobs = list(nhist, phist), nrow = 2, top = textGrob("Dissimilarity to the central tendency"))
```

### 2021-04-14 analysis/reports/submission2/table1_extreme_values.Rmd
maybe?

---
title: "Proportions of extreme values the unusual direction"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    # github_document:
    #    df_print: kable
    #    toc: true
    word_document:
        df_print: kable
        toc: false
---

>>>>

---
title: "Proportions of extreme values the usual direction"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
       toc: true
    word_document:
        df_print: kable
        toc: false
---
